### 2019-07-12 manuscript/methods.tex
Adding the correct figures and tables to results and first part of supplement

In order to test our hypotheses about lineage diversification and trait macroevolution, we fitted 29 state-dependent speciation and extinction models (SSE) based on previous work from \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
SSE models parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).

@@

We additionally fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
These models differ in whether transitions between the hidden states are symmetric or asymmetric, and whether the polyploidization rate depends on the hidden state.
Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of differences explained by an unknown factor.

We further fit the analogues of these five models but including a rate parameter for transitions from polyploid to diploid (M6--M10).
This allows us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.

@@

We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions).
Similar models were used by \citet{freyman_2017} used to study diversification in Onagraceae.

@@

As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by both ploidy and breeding system as well as a hidden trait (M18--M20, with varying degrees of complexity in the hidden trait transitions).
We add this hidden trait layer on top of our three-state model \citep[similar to][]{caetano_2018, huang_2018}.
We also fit the analogous models but allowing for diploidization (M21--M25).

@@

In order to compare model fits for ploidy only versus breeding system only versus both traits, we use the technique of \citet{tarasov_2019} for lumping states together.
We use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
First we lump together $ID$ and $CD$ to form the diploid state, mimicking the $D/P$ model (M26).
Then we lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
We further add a hidden trait to each of these models (M27 and M29).

@@

Parameters for each of the ten models were estimated using RevBayes \citep{hoehna_2016}.
Code for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the  models. % FIXME see referee comment about reporting convergence, not runtime
Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.

@@

We calculated the marginal likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).

@@

% \begin{table}
% \begin{tabular}{lcccccc} \toprule
% \begin{tabular}{@{}lcccccc@{}} \toprule
%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
%Model                     & Ploidy & Diploidization & Breeding  & Hidden & Num        & Marginal       \\
%                          &        &                & System    & State  & Parameters & Log-Likelihood \\ \midrule
%1. D/P                    & Yes    & No             & No        & No     & 5          & -1193.66\\
%2. D/P +$\delta$          & Yes    & Yes            & No        & No     & 6          & -1182.93 \\
%3. D/P+A/B                & Yes    & No             & No        & Yes    & 10         & -1150.99\\
%4. D/P+A/B+$\delta$       & Yes    & Yes            & No        & Yes    & 11         & \textbf{-1145.69}\\
%5. I/C                    & No     & No             & Yes       & No     & 5          & -1194.80 \\
%6. I/C+A/B                & No     & No             & Yes       & Yes    & 10         & \textbf{-1155.37}\\
%7. ID/CD/CP               & Yes    & No             & Yes       & No     & 9          & -1345.87\\
%8. ID/CD/CP+$\delta$      & Yes    & Yes            & Yes       & No     & 10         & -1344.50\\
%9. ID/CD/CP+A/B           & Yes    & No             & Yes       & Yes    & 15         & -1303.55 \\ 
%10. ID/CD/CP+A/B+$\delta$ & Yes    & Yes            & Yes       & Yes    & 16         & \textbf{-1300.35} \\
%\bottomrule
%\end{tabular}
%\caption{
   % The ten models and their marginal likelihoods.
    %Values in bold are for the best models within each class that are comparable (see \cref{table:bayesfactors}).
    %Abbreviations are D: diploid, P: polyploid, I: self-incompatible, C: self-compatible, A: one state of hidden trait, B: other state of hidden trait, $\delta$: diploidization.}
%\label{table:marginallike}
%\end{table}

\begin{table}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|c|c|c|}
\bottomrule
\end{tabular}
% E: We will probably have to change the layout here because it's not technically a table.  But good enough for now.
\caption{
    Bayes factors for model comparisons.
    Each of the three boxes contains models that can be compared with one another, based on the character states they include (see \cref{figure:stateclassifications}B).
    Models are numbered in as \cref{table:marginallike}. % FIXME: they aren't, but it's a pain to change; change in prev table instead?
    Bayes factors are reported on the natural log scale, so numbers greater than $+2$ mean that the model in the row label has `positive' support relative to the model in the column label; numbers less than $-2$ mean that model in the column label is the preferred one.
    Conventional thresholds for `strong' and `very strong' support are 6 and 10, respectively.
    The best model in each set is written in bold.
    In each case, it is the most complex model of the set.
}
\label{table:bayesfactors}
\end{table}

% E: Could we include a figure of the traits on the tree?  At least as supp info. FIXME

>>>>

In order to test our hypotheses about lineage diversification and trait macroevolution, we fitted 29 state-dependent speciation and extinction models (SSE) to previous work on diversification \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
SSE models parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).

@@

We additionally fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
These models differ in whether transitions between the hidden states are symmetric or asymmetric, and whether the polyploidization rate depends on the hidden state.
Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor\cref{table:bayesfactors}.

We further fit the analogues of these five models but including a rate parameter for transitions from polyploid to diploid (M6--M10) \cref{supptable:M6M10}.
This allows us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.

@@

We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions \cref{table:M11M15}).
Similar models were used by \citet{freyman_2017} used to study diversification in Onagraceae.

@@

As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by both ploidy and breeding system as well as a hidden trait (M18--M20, with varying degrees of complexity in the hidden trait transitions \cref{supptable:M16M20})
We add this hidden trait layer on top of our three-state model \citep[similar to][]{caetano_2018, huang_2018}.
We also fit the analogous models but allowing for diploidization (M21--M25, \cref{supptable:M21M25}).

@@

In order to compare model fits for ploidy only versus breeding system only versus both traits, we use the technique of \citet{tarasov_2019} for lumping states together.
We use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
First we lump together $ID$ and $CD$ to form the diploid state, mimicking the $D/P$ model (M26).
Then we lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
We further add a hidden trait to each of these models (M27 and M29).
We compared those models in table \cref{table:lumped}

@@

Parameters for each of the 29 models were estimated using RevBayes \citep{hoehna_2016}.
Code for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the  models. % FIXME see referee comment about reporting convergence, not runtime
Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.

@@

We calculated the marginal log-likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).

@@

% \begin{table}
% \begin{tabular}{lcccccc} \toprule
% \begin{tabular}{@{}lcccccc@{}} \toprule
%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
%Model                     & Ploidy & Diploidization & Breeding  & Hidden & Num        & Marginal       \\
%                          &        &                & System    & State  & Parameters & Log-Likelihood \\ \midrule
%1. D/P                    & Yes    & No             & No        & No     & 5          & -1193.66\\
%2. D/P +$\delta$          & Yes    & Yes            & No        & No     & 6          & -1182.93 \\
%3. D/P+A/B                & Yes    & No             & No        & Yes    & 10         & -1150.99\\
%4. D/P+A/B+$\delta$       & Yes    & Yes            & No        & Yes    & 11         & \textbf{-1145.69}\\
%5. I/C                    & No     & No             & Yes       & No     & 5          & -1194.80 \\
%6. I/C+A/B                & No     & No             & Yes       & Yes    & 10         & \textbf{-1155.37}\\
%7. ID/CD/CP               & Yes    & No             & Yes       & No     & 9          & -1345.87\\
%8. ID/CD/CP+$\delta$      & Yes    & Yes            & Yes       & No     & 10         & -1344.50\\
%9. ID/CD/CP+A/B           & Yes    & No             & Yes       & Yes    & 15         & -1303.55 \\ 
%10. ID/CD/CP+A/B+$\delta$ & Yes    & Yes            & Yes       & Yes    & 16         & \textbf{-1300.35} \\
%\bottomrule
%\end{tabular}
%\caption{
   % The ten models and their marginal likelihoods.
    %Values in bold are for the best models within each class that are comparable (see \cref{table:}).
    %Abbreviations are D: diploid, P: polyploid, I: self-incompatible, C: self-compatible, A: one state of hidden trait, B: other state of hidden trait, $\delta$: diploidization.}
%\label{table:marginallike}
%\end{table}

\begin{table}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|l|r|c|c|c|c|l|}
\hline
Model                & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M2    & M3     & M4    & M5     & Evidence                                                                                     \\ \hline
M1. D/P              & -1283.76                                                          & 59.90 & 49.23  & 60.48 & 58.82  & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M1\end{tabular}            \\ \hline
\textbf{M2. CID D/P} & \textbf{-1223.86}                                                 &       & -10.66 & 0.579 & -1.079 & \begin{tabular}[c]{@{}l@{}}Strong preference for \\ M2 over M1\end{tabular}              \\ \hline
M3. D/P+A/B          & -1234.52                                                          &       &        & 11.24 & 9.58   & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
\textbf{M4. D/P+A/B asym}     & \textbf{-1223.28}                                                          &       &        &       & -1.658 & \begin{tabular}[c]{@{}l@{}}Moderate evidence for\\ only asymmetric hidden rates\end{tabular} \\\hline
M5. D/P+A/B all asym & -1224.93                                                          &       &        &       &        &          \\ \hline                                                                                   
\end{tabular}
\caption{Bayes factors of  ploidy only models without diplodization. Numbers greater that 1 mean moderate support relative to the model in the row label.  Conventional threshold for `strong' support is 10. The best models are written in bold. }
\label{table: bayesfactors}
\end{table}

\begin{table}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|l|r|c|c|c|c|l|}
\hline
Model                      & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M12   & M13   & M14   & M15   & Evidence                                                                                       \\ \hline
M11. I/C                   & -1309.07                                                          & 41.13 & 38.59 & 61.34 & 61.40 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M111\end{tabular}            \\ \hline
M12. CID I/C               & -1267.93                                                          &       & -2.53 & 20.21 & 20.37 & \begin{tabular}[c]{@{}l@{}}Models with asymmetric rates\\ are preferred over  M12\end{tabular} \\ \hline
M13. I/C+A/B               & -1270.47                                                          &       &       & 22.75 & 22.80 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
\textbf{M14. I/C+A/B asym} & \textbf{-1247.72}                                                 &       &       &       & 0.05 & No evidence                                                                                    \\ \hline
\textbf{M15. I/C+A/B all asym }     & \textbf{ -1247.66}                                                          &       &       &       &       &       \\ \hline                                                                                        
\end{tabular}
\caption{Bayes factors of breeding system only models. Results indicate that the HiSSE models with asymmetric rates (M14, M15) are strongly preferred over every other model.}
\label{table:M11M15}
\end{table}

% E: Could we include a figure of the traits on the tree?  At least as supp info. FIXME
\begin{table}[]
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|l|c|c|c|c|}
\hline
Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=ln(BF(M1,M2)) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
M26. Lumped D/P             & -1463.22                                                          & M26 vs. M16 & 4.10            & M16 (Moderate)                                                       \\
\textbf{M16. ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
M27. Lumped D/P+A/B         & -1417.67                                                          & M27 vs. M23 & 3.678           & M23 (Moderate)                                                       \\
\textbf{M23. ID/CD/CP +A/B} & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
M28. Lumped I/C             & -1458.41                                                          & M28 vs. M16 & -0.69           & No evidence                                                          \\
\textbf{M16. ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
M29. Lumped I/C+ A/B        & -1416.60                                                          & M29 vs. M23 & 2.60            & M23 (Moderate)                                                       \\
\textbf{M23. IC/CD/CP+A/B}  & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
\end{tabular}
\caption{Test for adding a focal trait to a binary state model via Bayes factors. Three-state models are moderately preferred over two-state models. The evidence is stronger towards the inclusion of breeding systems in ploidy models (M26, M27) than the inclusion of ploidy in breeding system models (M29, M29)}
\label{table:lumped}
\end{table}
